Distributed Range-Only Localisation that Preserves Sensor and Navigator Privacies

### 2022-04-20 main.tex
Made all small changes from Benjamin's final proof read. Updated diff and response accordingly

\begin{abstract}
Distributed state estimation and localisation methods have become increasingly popular with the rise of ubiquitous computing, and have led naturally to an increased concern regarding data and estimation privacy. Traditional distributed sensor navigation methods typically involve the leakage of sensor or navigator information by communicating measurements or estimates and thus do not preserve participants' privacy. The existing approaches that do provide such guarantees, fail to address sensor and navigator privacy in the common application of model-based range-only estimation, consequently forfeiting broad applicability. In this work, we define a notion of privacy-preserving linear combination aggregation and use it to derive a modified Extended Kalman Filter using range measurements such that navigator location, sensors' locations, and sensors' measurements are kept private during navigation. Additionally, a formal cryptographic backing is presented to guarantee our method's privacy as well as an implementation to evaluate its performance. The novel, provably secure, range-based localisation method has applications in a variety of environments where sensors may not be trusted or estimates are considered sensitive, such as autonomous vehicle localisation or air traffic navigation.
\end{abstract}

@@

In section \ref{sec:problem_statement}, we introduce both the cryptographic and estimation problems considered in this work, before giving some relevant preliminaries in section \ref{sec:preliminaries}. Section \ref{sec:lcao_scheme} proposes a cryptographic scheme meeting our desired security properties and section \ref{sec:priv_localisation} gives a solution to the estimation problem making use of this scheme. Simulation results are discussed in section \ref{sec:sim_and_results} and concluding remarks are made in section \ref{sec:conclusion}.

@@

\subsection{Notation}
Throughout this work, we make use of the following notation. Underlined characters, $\vec{a}$, denote vectors, uppercase bold characters, $\mat{A}$, denote matrices. $\mat{A}^{-1}$ is the matrix inverse while $\mat{A}^\top$ is its transpose. The expected value of a random variable is denoted $\mathsf{E}[\cdot]$, while the variance of a random scalar and covariance of a random vector by $\mathsf{Var}[\cdot]$ and $\mathsf{Cov}[\cdot]$, respectively. $|a|$ denotes absolute value, $\lVert\vec{a}\rVert$ the vector norm, $\{\dots\}$ is used for sets and $\langle\dots\rangle$ for ordered sequences. When estimating a state, notation $\hat{\vec{x}}_{k|e}$ will denote an estimate of $\vec{x}$ at timestep $k$ given measurements from timesteps up to and including timestep $e$. When discussing cryptography, $\mathcal{E}_{sk}(\cdot)$ and $\mathcal{D}_{sk}(\cdot)$ will denote encryption and decryption with key $sk$, respectively, with $sk$ ommited when inferrable from context. $\mathbb{Z}_N$ is used for the set of integers modulo $N$ and $\mathbb{Z}^*_N$ for its multiplicative group. $\mathsf{lcm}(\cdot,\cdot)$ is the lowest common denominator, $\|$ the binary concatination operator and the term \textit{negligible function} refers to its cryptographic definition in \cite{katzIntroductionModernCryptography2008}.

@@

\section{Problem Statement} \label{sec:problem_statement}
In this work, we consider the context of privacy-preserving range sensor navigation, where we want no sensor to learn any information about the navigator or other sensors beyond their local measurements, and the navigator not to learn any information about individual sensors beyond its location estimate. The problem is two-fold, in that we require explicit cryptographic requirements with a suitable encryption scheme meeting them, as well as an estimation scheme that can use the encryption in the context of range-only navigation.

To give a formal cryptographic requirement in a distributed setting, we must first consider the communication requirements of our context, and define the attacker capabilities and the desired security of a suitable encryption scheme. In this section, we will define a communication protocol and the relevant formal definition of security we aim to achieve, followed by the estimation problem to which we will apply it.

@@

We are now ready to define the type of encryption scheme we want for the specified communication protocol and the security guarantees it should provide. We let a linear combination aggregation scheme be defined as a tuple of the four algorithms $(\mathsf{Setup}, \mathsf{Enc}, \mathsf{CombEnc}, \mathsf{AggDec})$. These will be used by a trusted setup party, the navigator and sensors $i\in\{1,\dots,n\}$. They are defined as follows.
\begin{LaTeXdescription}
    \item[$\mathsf{Setup}(\kappa)$] On input of security parameter $\kappa$, generate public parameters $\mathsf{pub}$, the number of weights $m$, the navigator's public and private keys $pk_0$ and $sk_0$ and the sensor private keys $sk_i,\,i\in\{1,\dots,n\}$.
    \item[$\mathsf{Enc}(pk_0, x)$] The navigator and sensors can encrypt any value $x$ with the navigator's public key $pk_0$ and obtain the encryption $\mathcal{E}_{pk_0}(x)$.
    \item[$\mathsf{CombEnc}(t, pk_0, sk_i, \mathcal{E}(\omega_1^{(t)}),\dots,\mathcal{E}(\omega_m^{(t)}), a^{(t)}_{i,1},\dots,a^{(t)}_{i,m})$] At instance $t$, sensor $i$ computes and obtains the encrypted linear combination denoted $l^{(t)}_i = \mathcal{E}_{pk_0,sk_i}(\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j)$ using its secret key $sk_i$.
    \item[$\mathsf{AggDec}(t, pk_0, sk_0, l^{(t)}_1,\dots,l^{(t)}_n)$] At instance $t$, the navigator computes the aggregation of linear combinations $\sum^{n}_{i=1}l_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} a^{(t)}_{i,j}\omega^{(t)}_j$ using its public and private keys $pk_0$, $sk_0$.
\end{LaTeXdescription}
The security notions we want these algorithms to meet reflect the previously stated estimation privacy goals. The navigator should learn no information from individual sensors while sensors should learn no information from the navigator or any other sensors. In the context of the introduced communication protocol, this can be summarised as the following notions.
\begin{LaTeXdescription}
    \item[Indistinguishable Weights] No colluding subset of sensors gains any new knowledge about the navigator weights $\omega^{(t)}_j,\,j\in\{1,\dots,m\}$ when receiving only their encryptions from the current and previous instances and having the ability to encrypt plaintexts of their choice.
    \item[Linear Combination Aggregator Obliviousness] No colluding subset \textit{excluding} the navigator gains additional information about the remaining sensor values to be weighted, $a^{(t)}_{i,j},\,j\in\{1,\dots,m\}$, where sensor $i$ is not colluding, given only encryptions of their linear combinations $l_i$ from the current and previous instances. Any colluding subset \textit{including} the navigator learns only the sum of all linear combinations weighted by weights of their choice, $\sum^{n}_{i=1}l_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} a^{(t)}_{i,j}\omega^{(t)}_j$.
\end{LaTeXdescription}
While indistinguishable weights can be achieved by encrypting weights with an encryption scheme meeting the notion of Indistinguishability under the Chosen Plaintext Attack (IND-CPA) \cite{katzIntroductionModernCryptography2008}, the novel notion of Linear Combination Aggregator Obliviousness (LCAO) has been formalised as a typical cryptographic game between attacker and challenger in appendix \ref{app:lcao}. Lastly, we conclude the cryptographic problem definition with the following important remark.
\begin{remark}
    A leakage function including weights from the navigator requires extra care to be taken when giving its definition. If an attacker compromises the navigator, they have control over the weights, and therefore the leakage function. We note that in the leakage function above, $\sum^n_{i=1}\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j$, an individual sum weighted by the same weight may be learnt by an attacker, \textit{e.g.}, $\sum^n_{i=1}a^{(t)}_{i,1}$ given weights $(1,0,\dots,0)$, but that individual sensor values $a^{(t)}_{i,j}$ remain private due to the assumption of a consistent broadcast.
\end{remark}

@@

\section{Private Linear Combination Aggregation Scheme} \label{sec:lcao_scheme}
In this section, we introduce an encryption scheme meeting the desired security properties in section \ref{subsec:crypto_problem}. The scheme is a combination of the Paillier and Joye-Libert schemes and provides encrypted weights meeting IND-CPA and encrypted aggregation meeting the notion of LCAO defined in section \ref{subsec:crypto_problem}. Similarly to its constituents, the scheme bases its security on the DCRA and, as with the Joye-Libert scheme, requires a trusted party for initial key generation and distribution.

@@

    \item[$\mathsf{CombEnc}(t_{k,\tau}, pk_0, sk_i, \mathcal{E}(\omega_1^{(k,\tau)})\dots , a^{(k,\tau)}_{i,1}\dots )$] At instance $t_{k,\tau}$, encrypted linear combination is computed as 
    \begin{equation}
        l^{(k,\tau)}_i = H(t_{k,\tau})^{sk_i}\prod^{m}_{j=1}\mathcal{E}_{pk_0}(\omega^{(k,\tau)}_j)^{a^{(k,\tau)}_{i,j}} \pmod{N^2}\,,\label{eqn:our_scheme_lin_comb}
    \end{equation}
    and makes use of the homomorphic property \eqref{eqn:paillier_hom_mult}. Correctness follows from
    \begin{equation*}
        \begin{split}
            l^{(k,\tau)}_i &= H(t_{k,\tau})^{sk_i}\prod^{m}_{j=1}\mathcal{E}_{pk_0}(\omega^{(k,\tau)}_j)^{a^{(k,\tau)}_{i,j}} \pmod{N^2} \\
            &= H(t_{k,\tau})^{sk_i}\prod^{m}_{j=1}\mathcal{E}_{pk_0}(a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j) \pmod{N^2} \\
            &= H(t_{k,\tau})^{sk_i}\cdot \\
            &\qquad \prod^{m}_{j=1}(N+1)^{a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j} \rho^{N}_{j} \pmod{N^2} \\
            &= H(t_{k,\tau})^{sk_i}\cdot \\
            &\qquad (N+1)^{\sum^{m}_{j=1}a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j} \rho_{i}^{N} \pmod{N^2}\,,
        \end{split}
    \end{equation*}
    for some values $\rho_i,\rho_j \in \mathbb{Z}_N,\,j\in\{1,\dots,m\}$. Here, $\rho_i^N$ and $H(t_{k,\tau})^{sk_i}$ can be considered the noise terms corresponding to the two levels of encryption from $pk_0$ and $sk_i$, respectively.

    \item[$\mathsf{AggDec}(t_{k,\tau}, pk_0, sk_0, l^{(k,\tau)}_1,\dots,l^{(k,\tau)}_n)$] Aggregation is computed as $l^{(k,\tau)} = \prod^n_{i=1}l^{(k,\tau)}_i \pmod{N^2}$, removing the aggregation noise terms, and is followed by Paillier scheme decryption
    \begin{equation}
        \begin{split}
            \sum^{n}_{i=1}\sum^{m}_{j=1}&a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j =\\
            &\frac{L((l^{(k,\tau)})^\lambda\pmod{N^2})}{L((N+1)^\lambda\pmod{N^2})} \pmod{N}\,,
        \end{split} \label{eqn:our_scheme_decrypt}
    \end{equation}
    with $\lambda = \mathsf{lcm}(p-1, q-1)$ and $L(\psi) = \frac{\psi-1}{N}$. The correctness of the aggregation can be seen from
    \begin{align*}
        \begin{split}
        l^{(k,\tau)} &= \prod^n_{i=1}H(t_{k,\tau})^{sk_i}\cdot \\
        &\qquad (N+1)^{\sum^{m}_{j=1}a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j}\rho_i^N \!\!\pmod{N^2}
        \end{split}\\
        \begin{split}
            &= H(t_{k,\tau})^{\sum^n_{i=1}sk_i}\cdot \\
            &\qquad \prod^n_{i=1}(N+1)^{\sum^{m}_{j=1}a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j}\rho_i^N \!\!\pmod{N^2}
        \end{split}\\
        &= (N+1)^{\sum^n_{i=1}\sum^{m}_{j=1}a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j}\rho'^N \!\!\pmod{N^2}\,,
    \end{align*}
    for some values $\rho_i,\rho' \in \mathbb{Z}_N,\,i\in\{1,\dots,n\}$.
\end{LaTeXdescription}
Additionally, we note that in the above construction, all weights $\omega^{(k,\tau)}_j$ and values $a^{(k,\tau)}_{i,j}$ are integers and the resulting linear combinations and aggregation are computed modulo $N$.

@@

\section{Privacy-Preserving Localisation} \label{sec:priv_localisation}
With a concrete scheme meeting the LCAO notion, we can now put forward a localisation filter with communication that can be reformulated to the required protocol. To produce an estimate of the state $\vec{x}_k$, we make use of an algebraic reformulation of the Extended Kalman Filter (EKF), the Extended Information Filter (EIF) \cite{maybeckStochasticModelsEstimation1982}, which reduces the filter update step to a single summation. The EIF update step requires the predicted state estimate $\hat{\vec{x}}_{k|k-1}$ and estimate covariance $\mat{P}_{k|k-1}$ in the information vector and matrix forms
\begin{equation}
    \hat{\vec{y}}_{k|k-1} = \mat{P}_{k|k-1}^{-1}\hat{\vec{x}}_{k|k-1} \quad\textrm{and}\quad \mat{Y}_{k|k-1} = \mat{P}_{k|k-1}^{-1}\,,
\end{equation}
respectively. In this form, the update equations for $n$ sensor measurements at time $k$, with measurement models \eqref{eqn:measurement_model}, are given by
\begin{equation}
    \begin{split}
        &\hat{\vec{y}}_{k|k} = \hat{\vec{y}}_{k|k-1} + \\
        &\quad \sum^n_{i=1}\mat{H}^\top_{k,i} r^{-1}_i (z_{k,i} - h_i(\hat{\vec{x}}_{k|k-1}) + \mat{H}_{k,i}\hat{\vec{x}}_{k|k-1})
    \end{split} \label{eqn:eif_info_vec_update}
\end{equation}
and
\begin{equation}
    \mat{Y}_{k|k} = \mat{Y}_{k|k-1} + \sum^n_{i=1}\mat{H}^\top_{k,i} r^{-1}_i \mat{H}_{k,i}\,, \label{eqn:eif_info_mat_update}
\end{equation}
with Jacobians
\begin{equation}
    \mat{H}_{k,i} = \left.\frac{\partial h_i}{\partial \vec{x}}\right|_{\hat{\vec{x}}_{k|k-1}} \label{eqn:measurement_jacobian}
\end{equation}
for sensors $i\in\{1,\dots,n\}$. After converting the updated information vector and matrix back to state estimate $\hat{\vec{x}}_{k|k}$ and estimate covariance $\mat{P}_{k|k}$, the filter's prediction step can be computed by the navigator locally using any suitable filter for the known system model \eqref{eqn:system_model}.

In the form above, at every timestep $k$, all sensitive sensor information required for state estimation is captured in the measurement vector
\begin{equation}
    \vec{i}_{k,i} = \mat{H}^\top_{k,i} r^{-1}_i (z_{k,i} - h_i(\hat{\vec{x}}_{k|k-1}) + \mat{H}_{k,i}\hat{\vec{x}}_{k|k-1}) \label{eqn:measurement_vec}
\end{equation}
and the measurement matrix
\begin{equation}
    \mat{I}_{k,i} = \mat{H}^\top_{k,i} r^{-1}_i \mat{H}_{k,i}\,, \label{eqn:measurement_mat}
\end{equation}
namely, their measurements $z_{k,i}$, measurement variances $r_{k,i}$ and locations $\vec{s}_i$; captured in measurement functions $h_i$ and Jacobians $\mat{H}_{k,i}$. To compute $\vec{i}_{k,i}$ and $\mat{I}_{k,i}$, however, the current predicted state estimate $\hat{\vec{x}}_{k|k-1}$ is also required (in $h_i$ and $\mat{H}_{k,i}$). Therefore, our goal is to rearrange \eqref{eqn:measurement_vec} and \eqref{eqn:measurement_mat} as a linear combination of functions of $\hat{\vec{x}}_{k|k-1}$ (the navigator weights) computable at each sensor $i$, to be subsequently aggregated at the navigator. Application of the linear combination aggregation scheme proposed in turn guarantees that sensors do not learn the navigator state, and the navigator learns only the aggregation required for updating its state estimate in \eqref{eqn:eif_info_vec_update} and \eqref{eqn:eif_info_mat_update}.

@@

The above rearrangements give $\vec{i}'_{k,i}$ and $\mat{I}'_{k,i}$ as linear comabinations of elements in

@@

            \item[\textit{Combine Query}] When $\mathcal{A}$ submits combine query $(i, t, a^{(t)}_{i,1},\dots,a^{(t)}_{i,m})$, choose weights $\omega^{(t)}_j,j\in\{1,\dots,m\}$ if not already chosen for the instance $t$, and make an AO encryption query $(i, t, \sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j)$ to $\mathcal{C}$. The received response will be of the form $(N+1)^{\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j}H(t)^{sk_i}$; multiply it by $\rho^N$ for a random $\rho \in \mathbb{Z}_N$ and return 

@@

                (N+1)^{\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j}\rho^N H(t)^{sk_i} \pmod{N^2}

@@

        for an unknown $b \in \{0,1\}$. Multiply series elements by $\rho_i^N,\,i\in\{1,\dots,n\}$ for randomly chosen $\rho_i \in \mathbb{Z}_N$ and return

>>>>

\begin{abstract}
Distributed state estimation and localisation methods have become increasingly popular with the rise of ubiquitous computing, and have led naturally to an increased concern regarding data and estimation privacy. Traditional distributed sensor navigation methods typically involve the leakage of sensor or navigator information by communicating measurements or estimates and thus do not preserve participants' privacy. The existing approaches that do provide such guarantees fail to address sensor and navigator privacy in the common application of model-based range-only localisation, consequently forfeiting broad applicability. In this work, we define a notion of privacy-preserving linear combination aggregation and use it to derive a modified Extended Kalman Filter using range measurements such that navigator location, sensors' locations, and sensors' measurements are kept private during navigation. Additionally, a formal cryptographic backing is presented to guarantee our method's privacy as well as an implementation to evaluate its performance. The novel, provably secure, range-based localisation method has applications in a variety of environments where sensors may not be trusted or estimates are considered sensitive, such as autonomous vehicle localisation or air traffic navigation.
\end{abstract}

@@

In section \ref{sec:problem_statement}, we introduce both the cryptographic and estimation problems considered in this work, before giving some relevant preliminaries in section \ref{sec:preliminaries}. Section \ref{sec:lcao_scheme} proposes a cryptographic scheme meeting our desired security properties, and section \ref{sec:priv_localisation} gives a solution to the estimation problem making use of this scheme. Simulation results are discussed in section \ref{sec:sim_and_results}, and concluding remarks are made in section \ref{sec:conclusion}.

@@

\subsection{Notation}
Throughout this work, we make use of the following notation. Underlined characters, $\vec{a}$, denote vectors, uppercase bold characters, $\mat{A}$, denote matrices. $\mat{A}^{-1}$ is the matrix inverse while $\mat{A}^\top$ is its transpose. The expected value of a random variable is denoted by $\mathsf{E}[\cdot]$, while the variance of a random scalar and covariance of a random vector by $\mathsf{Var}[\cdot]$ and $\mathsf{Cov}[\cdot]$, respectively. $|a|$ denotes absolute value, $\lVert\vec{a}\rVert$ the vector norm, $\{\dots\}$ is used for sets and $\langle\dots\rangle$ for ordered sequences. When estimating a state, notation $\hat{\vec{x}}_{k|e}$ will denote an estimate of $\vec{x}$ at timestep $k$ given measurements from timesteps up to and including timestep $e$. When discussing cryptography, $\mathcal{E}_{sk}(\cdot)$ and $\mathcal{D}_{sk}(\cdot)$ will denote encryption and decryption with key $sk$, respectively, with $sk$ ommited when inferrable from context. $\mathbb{Z}_N$ is used for the set of integers modulo $N$ and $\mathbb{Z}^*_N$ for its multiplicative group. $\mathsf{lcm}(\cdot,\cdot)$ is the lowest common denominator, $\|$ the binary concatination operator, and the term \textit{negligible function} refers to its cryptographic definition in \cite{katzIntroductionModernCryptography2008}.

@@

\section{Problem Statement} \label{sec:problem_statement}
In this work, we consider the context of privacy-preserving range sensor navigation, where we want no sensor to learn any information about the navigator or other sensors beyond their local measurements, and the navigator not to learn any information about individual sensors beyond its location estimate. The problem is two-fold, in that we require explicit cryptographic requirements with a suitable encryption scheme meeting them as well as an estimation scheme that can use the encryption in the context of range-only navigation.

To give a formal cryptographic requirement in a distributed setting, we must first consider the communication requirements of our context and define the attacker capabilities and the desired security of a suitable encryption scheme. In this section, we will define a communication protocol and the relevant formal definition of security we aim to achieve, followed by the estimation problem to which we will apply it.

@@

We are now ready to define the type of encryption scheme we want for the specified communication protocol and the security guarantees it should provide. We let a linear combination aggregation scheme be defined as a tuple of the four algorithms $(\mathsf{Setup}, \mathsf{Enc}, \mathsf{CombEnc}, \mathsf{AggDec})$. These will be used by a trusted setup party, the navigator, and sensors $i\in\{1,\dots,n\}$. They are defined as follows.
\begin{LaTeXdescription}
    \item[$\mathsf{Setup}(\kappa)$] On input of security parameter $\kappa$, generate public parameters $\mathsf{pub}$, the number of weights $m$, the navigator's public and private keys $pk_0$ and $sk_0$ and the sensor private keys $sk_i,\,i\in\{1,\dots,n\}$.
    \item[$\mathsf{Enc}(pk_0, x)$] The navigator and sensors can encrypt any value $x$ with the navigator's public key $pk_0$ and obtain the encryption $\mathcal{E}_{pk_0}(x)$.
    \item[$\mathsf{CombEnc}(t, pk_0, sk_i, \mathcal{E}(\omega_1^{(t)}),\dots,\mathcal{E}(\omega_m^{(t)}), a^{(t)}_{i,1},\dots,a^{(t)}_{i,m})$] At instance $t$, sensor $i$ computes and obtains the encrypted linear combination denoted $l^{(t)}_i = \mathcal{E}_{pk_0,sk_i}(\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j)$ using its secret key $sk_i$.
    \item[$\mathsf{AggDec}(t, pk_0, sk_0, l^{(t)}_1,\dots,l^{(t)}_n)$] At instance $t$, the navigator computes the aggregation of linear combinations $\sum^{n}_{i=1}l_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} a^{(t)}_{i,j}\omega^{(t)}_j$ using its public and private keys $pk_0$, $sk_0$.
\end{LaTeXdescription}
The security notions we want these algorithms to meet reflect the previously stated estimation privacy goals. The navigator should learn no information from individual sensors while sensors should learn no information from the navigator or any other sensors. In the context of the introduced communication protocol, this can be summarised as the following notions.
\begin{LaTeXdescription}
    \item[Indistinguishable Weights] No colluding subset of sensors gains any new knowledge about the navigator weights $\omega^{(t)}_j,\,j\in\{1,\dots,m\}$ when receiving only their encryptions from the current and previous instances and having the ability to encrypt plaintexts of their choice.
    \item[Linear Combination Aggregator Obliviousness] No colluding subset \textit{excluding} the navigator gains additional information about the remaining sensor values to be weighted, $a^{(t)}_{i,j},\,j\in\{1,\dots,m\}$, where sensor $i$ is not colluding, given only encryptions of their linear combinations $l_i$ from the current and previous instances. Any colluding subset \textit{including} the navigator learns only the sum of all linear combinations weighted by weights of their choice, $\sum^{n}_{i=1}l_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} a^{(t)}_{i,j}\omega^{(t)}_j$.
\end{LaTeXdescription}
While indistinguishable weights can be achieved by encrypting weights with an encryption scheme meeting the notion of Indistinguishability under the Chosen Plaintext Attack (IND-CPA) \cite{katzIntroductionModernCryptography2008}, the novel notion of Linear Combination Aggregator Obliviousness (LCAO) has been formalised as a typical cryptographic game between attacker and challenger in appendix~\ref{app:lcao}. Lastly, we conclude the cryptographic problem definition with the following important remark.
\begin{remark}
    A leakage function including weights from the navigator requires extra care to be taken when giving its definition. If an attacker compromises the navigator, they have control over the weights, and therefore the leakage function. We note that in the leakage function above, $\sum^n_{i=1}\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j$, an individual sum weighted by the same weight may be learnt by an attacker, \textit{e.g.}, $\sum^n_{i=1}a^{(t)}_{i,1}$ given weights $(1,0,\dots,0)$, but that individual sensor values $a^{(t)}_{i,j}$ remain private due to the assumption of a consistent broadcast.
\end{remark}

@@

\section{Private Linear Combination Aggregation Scheme} \label{sec:lcao_scheme}
In this section, we introduce an encryption scheme meeting the desired security properties in section \ref{subsec:crypto_problem}. The scheme is a combination of the Paillier and Joye-Libert schemes and provides encrypted weights meeting IND-CPA and encrypted aggregation meeting the notion of LCAO defined in section~\ref{subsec:crypto_problem}. Similarly to its constituents, the scheme bases its security on the DCRA and, as with the Joye-Libert scheme, requires a trusted party for initial key generation and distribution.

@@

    \item[$\mathsf{CombEnc}(t_{k,\tau}, pk_0, sk_i, \mathcal{E}_{pk_0}(\omega_1^{(k,\tau)})\dots , a^{(k,\tau)}_{i,1}\dots )$] At the instance $t_{k,\tau}$, encrypted linear combination is given by 
    \begin{equation}
        l^{(k,\tau)}_i = H(t_{k,\tau})^{sk_i}\prod^{m}_{j=1}\mathcal{E}_{pk_0}(\omega^{(k,\tau)}_j)^{a^{(k,\tau)}_{i,j}} \pmod{N^2}\,,\label{eqn:our_scheme_lin_comb}
    \end{equation}
    and makes use of the homomorphic property \eqref{eqn:paillier_hom_mult}. Correctness follows from
    \begin{equation*}
        \begin{split}
            l^{(k,\tau)}_i &= H(t_{k,\tau})^{sk_i}\prod^{m}_{j=1}\mathcal{E}_{pk_0}(\omega^{(k,\tau)}_j)^{a^{(k,\tau)}_{i,j}} \pmod{N^2} \\
            &= H(t_{k,\tau})^{sk_i}\prod^{m}_{j=1}\mathcal{E}_{pk_0}(a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j) \pmod{N^2} \\
            &= H(t_{k,\tau})^{sk_i}\cdot \\
            &\qquad \prod^{m}_{j=1}(N+1)^{a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j} \rho^{N}_{j} \pmod{N^2} \\
            &= H(t_{k,\tau})^{sk_i}\cdot \\
            &\qquad (N+1)^{\sum^{m}_{j=1}a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j} \tilde{\rho}_{i}^{N} \pmod{N^2}\,,
        \end{split}
    \end{equation*}
    for some values $\rho_j \in \mathbb{Z}_N,\,j\in\{1,\dots,m\}$ and $\tilde{\rho}_i=\prod^{m}_{j=1}\rho_j$. Here, $\tilde{\rho}_i^N$ and $H(t_{k,\tau})^{sk_i}$ can be considered the noise terms corresponding to the two levels of encryption from $pk_0$ and $sk_i$, respectively.

    \item[$\mathsf{AggDec}(t_{k,\tau}, pk_0, sk_0, l^{(k,\tau)}_1,\dots,l^{(k,\tau)}_n)$] Aggregation is computed as $l^{(k,\tau)} = \prod^n_{i=1}l^{(k,\tau)}_i \pmod{N^2}$, removing the aggregation noise terms, and is followed by Paillier scheme decryption
    \begin{equation}
        \begin{split}
            \sum^{n}_{i=1}\sum^{m}_{j=1}&a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j =\\
            &\frac{L((l^{(k,\tau)})^\lambda\pmod{N^2})}{L((N+1)^\lambda\pmod{N^2})} \pmod{N}\,,
        \end{split} \label{eqn:our_scheme_decrypt}
    \end{equation}
    with $\lambda = \mathsf{lcm}(p-1, q-1)$ and $L(\psi) = \frac{\psi-1}{N}$. The correctness of the aggregation can be seen from
    \begin{align*}
        \begin{split}
        l^{(k,\tau)} &= \prod^n_{i=1}H(t_{k,\tau})^{sk_i}\cdot \\
        &\qquad (N+1)^{\sum^{m}_{j=1}a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j}\tilde{\rho}_i^N \!\!\pmod{N^2}
        \end{split}\\
        \begin{split}
            &= H(t_{k,\tau})^{\sum^n_{i=1}sk_i}\cdot \\
            &\qquad \prod^n_{i=1}(N+1)^{\sum^{m}_{j=1}a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j}\tilde{\rho}_i^N \!\!\pmod{N^2}
        \end{split}\\
        &= (N+1)^{\sum^n_{i=1}\sum^{m}_{j=1}a^{(k,\tau)}_{i,j}\omega^{(k,\tau)}_j}\tilde{\rho}^N \!\!\pmod{N^2}\,,
    \end{align*}
    for some values $\tilde{\rho}_i \in \mathbb{Z}_N,\,i\in\{1,\dots,n\}$ and $\tilde{\rho}=\prod^{n}_{i=1}\tilde{\rho}_i$.
\end{LaTeXdescription}
Additionally, we note that in the above construction, all weights $\omega^{(k,\tau)}_j$ and values $a^{(k,\tau)}_{i,j}$ are integers and the resulting linear combinations and aggregation are computed modulo $N$.

@@

\section{Privacy-Preserving Localisation} \label{sec:priv_localisation}
With a concrete scheme meeting the LCAO notion, we can now put forward a localisation filter with communication that can be reformulated to the required protocol. To produce an estimate of the state $\vec{x}_k$, we make use of an algebraic reformulation of the Extended Kalman Filter (EKF), the Extended Information Filter (EIF) \cite{maybeckStochasticModelsEstimation1982}, which reduces the filter update step to a single summation. The EIF update step requires the predicted state estimate $\hat{\vec{x}}_{k|k-1}$ and estimate covariance $\mat{P}_{k|k-1}$ in the information vector and matrix forms
\begin{equation}
    \hat{\vec{y}}_{k|k-1} = \mat{P}_{k|k-1}^{-1}\hat{\vec{x}}_{k|k-1} \quad\textrm{and}\quad \mat{Y}_{k|k-1} = \mat{P}_{k|k-1}^{-1}\,,
\end{equation}
respectively. In this form, the update equations for $n$ sensor measurements at time $k$, with measurement models \eqref{eqn:measurement_model}, are given by
\begin{equation}
    \begin{split}
        &\hat{\vec{y}}_{k|k} = \hat{\vec{y}}_{k|k-1} + \\
        &\quad \sum^n_{i=1}\mat{H}^\top_{k,i} r^{-1}_i \left(z_{k,i} - h_i(\hat{\vec{x}}_{k|k-1}) + \mat{H}_{k,i}\hat{\vec{x}}_{k|k-1}\right)
    \end{split} \label{eqn:eif_info_vec_update}
\end{equation}
and
\begin{equation}
    \mat{Y}_{k|k} = \mat{Y}_{k|k-1} + \sum^n_{i=1}\mat{H}^\top_{k,i} r^{-1}_i \mat{H}_{k,i}\,, \label{eqn:eif_info_mat_update}
\end{equation}
with Jacobians
\begin{equation}
    \mat{H}_{k,i} = \left.\frac{\partial h_i}{\partial \vec{x}}\right|_{\hat{\vec{x}}_{k|k-1}} \label{eqn:measurement_jacobian}
\end{equation}
for sensors $i\in\{1,\dots,n\}$. After converting the updated information vector and matrix back to state estimate $\hat{\vec{x}}_{k|k}$ and estimate covariance $\mat{P}_{k|k}$, the filter's prediction step can be computed by the navigator locally using any suitable filter for the known system model \eqref{eqn:system_model}.

In the form above, at every timestep $k$, all sensitive sensor information required for state estimation is captured in the measurement vector
\begin{equation}
    \vec{i}_{k,i} = \mat{H}^\top_{k,i} r^{-1}_i \left(z_{k,i} - h_i(\hat{\vec{x}}_{k|k-1}) + \mat{H}_{k,i}\hat{\vec{x}}_{k|k-1}\right) \label{eqn:measurement_vec}
\end{equation}
and the measurement matrix
\begin{equation}
    \mat{I}_{k,i} = \mat{H}^\top_{k,i} r^{-1}_i \mat{H}_{k,i}\,, \label{eqn:measurement_mat}
\end{equation}
namely, their measurements $z_{k,i}$, measurement variances $r_{k,i}$ and locations $\vec{s}_i$; captured in measurement functions $h_i$ and Jacobians $\mat{H}_{k,i}$. To compute $\vec{i}_{k,i}$ and $\mat{I}_{k,i}$, however, the current predicted state estimate $\hat{\vec{x}}_{k|k-1}$ is also required (in $h_i$ and $\mat{H}_{k,i}$). Therefore, our goal is to rearrange \eqref{eqn:measurement_vec} and \eqref{eqn:measurement_mat} as a linear combination of functions of $\hat{\vec{x}}_{k|k-1}$ (the navigator weights) computable at each sensor $i$, to be subsequently aggregated at the navigator. Application of the linear combination aggregation scheme proposed in turn guarantees that sensors do not learn the navigator state, and the navigator learns only the aggregation required for updating its state estimate in \eqref{eqn:eif_info_vec_update} and \eqref{eqn:eif_info_mat_update}.

@@

The above rearrangements give $\vec{i}'_{k,i}$ and $\mat{I}'_{k,i}$ as linear combinations of elements in

@@

            \item[\textit{Combine Query}] When $\mathcal{A}$ submits combine query $(i, t, a^{(t)}_{i,1},\dots,a^{(t)}_{i,m})$, choose weights $\omega^{(t)}_j,j\in\{1,\dots,m\}$ if not already chosen for the instance $t$, and make an AO encryption query $(i, t, \sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j)$ to $\mathcal{C}$. The received response will be of the form $(N+1)^{\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j}H(t)^{sk_i}$; multiply it by $\tilde{\rho}^N$ for a random $\tilde{\rho} \in \mathbb{Z}_N$ and return 

@@

                (N+1)^{\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j}\tilde{\rho}^N H(t)^{sk_i} \pmod{N^2}

@@

        for an unknown $b \in \{0,1\}$. Multiply series elements by $\tilde{\rho}_i^N,\,i\in\{1,\dots,n\}$ for randomly chosen $\tilde{\rho}_i \in \mathbb{Z}_N$ and return
